---
title: "Fars"
author: "jluthy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r, include = FALSE}
library(dpylr)
library(readr)
library(tidyr)
library(maps)
```


```{r fars_functions, include = FALSE}
fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}

make_filename <- function(year) {
  year <- as.integer(year)
  sprintf("accident_%d.csv.bz2", year)
}

fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = year) %>%
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}

fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>%
    dplyr::group_by(year, MONTH) %>%
    dplyr::summarize(n = n()) %>%
    tidyr::spread(year, n)
}

fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)

  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}
```

This describes some of the features of this package. 
Reads data from FARS, a nationwide database of reported wrecks and fatalitits on public roads. 

#### *Read in the data*

Code was built to look through three of the provided data sets and return the matching file. It will then read the file and return a table.

```{r read data file}
fars_read(filename="accident_2015.csv.bz2")
```

#### *Create a file name*
This function creates the file name when given 'year' as input.

This function returns a character string with coding "accident_year.csv.bz2"

```{r make file name}
make_filename(2015)
```

#### *Read Data & Summarize*

This function makes data frames with MONTH and year columns

When passing the function a list of valid years as input, it gets associated files, For each file it retrieves the corresponding data table and selects the MONTH and year columns.

This function returns a list of tables corresponding to list of years.

```{r summary}
fars_read_years(years=c(2013,2014,2015))
```

#### *Summarize Years*

This function receives a list of years as an input and gets data tables with \code{fars_read_years} Then it summarizes the results per year.

```{r summarize years}
fars_summarize_years(years=c(2013,2014,2015))
```


#### *Map it out*

This function gets passed a specific state and year, then gets associated data. If the state exists, it filters the data per state and plots accidents for that state.

```{r map it}
fars_map_state(state.num=1,year=2015)
```
