---
title: "fars"
author: "jluthy"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r libs, include = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(maps)
```


This describes some of the features of this package. 
Reads data from FARS, a nationwide database of reported wrecks and fatalitits on public roads. 

### *Read in the data*

Code was built to look through three of the provided data sets and return the matching file. It will then read the file and return a table.

```{r read data file, inlcude = FALSE, eval = FALSE}
library(fars)
fars_read("accident_2015.csv.bz2")
```

### *Create a file name*
This function creates the file name when given year as input.

This function returns a character string with coding "accident_year.csv.bz2"

```{r make file name, inlcude = FALSE, eval = FALSE}
make_filename(c(2013, 2014, 2015))
```

### *Read Data & Summarize*

This function makes data frames with MONTH and year columns

When passing the function a list of valid years as input, it gets associated files, For each file it retrieves the corresponding data table and selects the MONTH and year columns.

This function returns a list of tables corresponding to list of years.

```{r summary, include = FALSE, eval = FALSE}
fars_read_years(c(2013,2014,2015))
```

### *Summarize Years*

This function receives a list of years as an input and gets data tables with \code{fars_read_years} Then it summarizes the results per year.

```{r summarize years, include = FALSE, eval = FALSE}
fars_summarize_years(c(2013,2014,2015))
```


### *Map it out*

This function gets passed a specific state and year, then gets associated data. If the state exists, it filters the data per state and plots accidents for that state.

```{r map it, include = FALSE, eval = FALSE}
fars_map_state(state.num=1,year=2015)
```
